Poster: Online RL in the Programmable Dataplane with \approachshort{}

### 2021-10-12 paper/poster-abstract.tex
Poster-er, with Reward lookup/storage fixes.
Not fixed on Ind yet, that one's gonna bite me in the proverbial, I assume...

% !TeX document-id = {5530719d-34df-4dd8-b4b5-e6ed092c2b36}
% !TeX program = pdflatex
% !BIB program = biber
\documentclass[
sigconf,natbib=false
%,anonymous=true
]{acmart}

@@

%\DeclareFieldFormat[inproceedings]{doi}{}
%\DeclareFieldFormat[article]{doi}{}
%\DeclareFieldFormat*{url}{}
%\DeclareFieldFormat[online]{url}{\mkbibacro{URL}\addcolon\space\url{#1}}
%\DeclareFieldFormat[report]{url}{\mkbibacro{URL}\addcolon\space\url{#1}}

@@

\DefineBibliographyStrings{english}{%
	andothers = {\emph{et al}\adddot}
}
\DeclareFieldFormat[inproceedings]{url}{}
\DeclareFieldFormat[article]{url}{}
%\DeclareFieldFormat[inproceedings]{doi}{}
%\DeclareFieldFormat[article]{doi}{}
%\DeclareFieldFormat[inproceedings]{editor}{}
%\DeclareFieldFormat[proceedings]{editor}{}
%\DeclareFieldFormat[article]{editor}{}

@@

%\vspace{-1em}
\section{Introduction}
\begin{figure}
	\centering
		\includegraphics[keepaspectratio, width=0.6875\linewidth]{figures/arch-with-p4}
		\caption{\approachshort{} brings low-latency, online reinforcement learning to SoC- and NPU-based SmartNICs. Classical RL policy methods are the key to making this computationally feasible.\label{fig:netro-arch}}
\end{figure}
\begin{figure}
	\centering
%	\includegraphics[keepaspectratio, width=0.6875\linewidth]{figures/arch-with-p4}
%	\caption{\approachshort{} brings low-latency, online reinforcement learning directly to the dataplane. SoC- and NetFPGA-based SmartNIC devices expose spare compute---making in-situ, asynchronous processing and learning possible alongside P4 dataplanes. Classical RL policy methods are the key to making this computationally feasible.\label{fig:netro-arch}}
	\centering
	\includegraphics[keepaspectratio, width=0.8\linewidth]{figures/coop}
	\caption{Parallel policy execution. A single \emph{controller} delegates RL computation and updates to many \emph{minion} threads, using a shared atomic writeback.\label{fig:single-and-parallel:parallel}}
\end{figure}
Automatic network optimisation, control, and defence are at last becoming commonplace through data-driven techniques such as RL methods, where every change and its effects improve future decisions made by an agent.
In tandem, P4~\parencite{DBLP:journals/ccr/BosshartDGIMRSTVVW14} and \emph{programmable dataplane} (PDP) hardware have inspired explosive growth and interest in the research community surrounding in-network computation and offloading.
This ecosystem exposes not only runtime reconfigurable packet processing, but per-packet and per-flow traffic measurement state that can greatly aid network operation~\parencite{DBLP:conf/sosr/GhasemiBR17}.

@@

%?? Do we want concrete examples of things `pushed down the stack'?
%
%?? Consider a DDoS prevention system...
%?? in the face of new data, Able to start learning again in microseconds
\fakepara{Machine learning in the dataplane}
To process this fine-grained state both at line rate \emph{and} at low latency, there has been keen interest in executing ML in the dataplane~\parencite{DBLP:conf/hotnets/XiongZ19,DBLP:journals/corr/abs-2009-02353}.
All works in this category, while powerful, operate by converting a \emph{pre-trained} model into a suitable representation, such as a binary neural network or string of match-action tables.
%These works have shown the value of in-network ML: high-throughput, low latency response to network changes.
These can react to on-device state, yet the missing piece of the puzzle is learning and updating these ML analyses \emph{online}, without deferring to another machine in the network.
%Training these models online and in-network is an exciting (and challenging) lacuna in the field that \emph{has yet to be addressed by the community}.
This lacuna has yet to be addressed by the community.

\fakepara{Problems of online training}
Programmable dataplane hardware, being designed solely for efficient packet processing, lacks floating-point arithmetic support even in the case of more general purpose NPU-type SmartNICs.
Additionally, deep neural network training relies on backpropagation, can require memorizing many a sizeable replay buffer for RL, and needs many mini-batches of data for stable training.
The natural solution is to pass training data to a host machine or dedicated accelerator, yet this adds delays in crossing the PCIe bus, kernel-user handover, and buffering---adding latency while learning.
The above (NIC-suitable) inference schemes still, even if their data formats solve the issue of floating point use.
Online learning thus requires careful choices in both sample data formats, and the learned function approximation scheme.

\section{\approachshort{}'s design}
\approachshort{} makes RL workloads feasible under on SmartNICs by using quantised fixed-point representations of action values, tile-coded policies, and one-step temporal-difference RL algorithms such as Sarsa~\parencite{RL2E}.
These allow us to evaluate and update policies using \emph{only integer arithmetic}.
We observe that tile coded inference is a map-reduce problem (\cref{fig:par-tilecode}), and over integers admits a novel \emph{wait-free Sarsa} RL implementation (exploiting the parallelism built in to SmartNICs).
In addition to their runtime simplicity, while these functions have lower model capacity they do not require batches of inputs to learn in a stable way, negating the memory needed to store experience replays.

@@

\fakepara{Inference/Updates}
State is mirrored to a set of packet processing threads, which compute hit tiles from a locally held work set.
During inference, these threads atomically add values to a shared action preference list, while policy updates are guaranteed by tile-coding to proceed without data races (\cref{fig:single-and-parallel:parallel}).
In the case of our Netronome implementation, these run on lightweight cores known as \emph{microengines} (MEs)---which could be substituted for suitable FUs on CGRA/FPGA hardware.
Additionally, we have a single-threaded implementation, \indfw, which forgoes this cooperative model to maximise offline throughput (i.e., by having many parallel agents handle state).

@@

\section{Evaluation}
We compare our implementation of \approachshort{} against a tile-coded RL agent written in numpy on a powerful host (i7-6700K@\qtyproduct[product-units=single]{4 x 4.2}{\giga\hertz}), using policy dimensions used by existing DDoS prevention work\parencite{DBLP:journals/tnsm/SimpsonRP20}, across various quantisation bit depths.
\Cref{fig:lat-cumul} shows that our main design (\coopfw) achieves order-of-magnitude improvements over our single-threaded variant (\indfw) and a host implementation, as well as considerably tighter tail behaviour on the SmartNIC.
Note that lower bit depths have higher execution costs due to the native \qty{32}{\bit} width of Netronome registers.
On update times, we see \qty{63.12}{\micro\second} at 99.99\nthscript{th} percentile for a \qty{32}{\bit} policy of this size.
We relate a \qty{15}{$\times$} latency (\coopfw) and \qty{2.82}{$\times$} offline throughput-per-core (\indfw) in spite of the Netronome's considerably slower clock speed (\qty{0.29}{$\times$}), showing the value of exploiting SmartNIC parallelism.

>>>>

% !TeX document-id = {5530719d-34df-4dd8-b4b5-e6ed092c2b36}
% !TeX program = pdflatex
% !BIB program = biber
\documentclass[
sigconf,natbib=false
,anonymous=true
]{acmart}

@@

\DeclareFieldFormat[inproceedings]{doi}{}
\DeclareFieldFormat[article]{doi}{}
\DeclareFieldFormat*{url}{}
%\DeclareFieldFormat[online]{url}{\mkbibacro{URL}\addcolon\space\url{#1}}
%\DeclareFieldFormat[report]{url}{\mkbibacro{URL}\addcolon\space\url{#1}}

@@

\DefineBibliographyStrings{english}{%
	andothers = {\emph{et al}\adddot}
}
\DeclareFieldFormat[inproceedings]{url}{}
\DeclareFieldFormat[article]{url}{}
%\DeclareFieldFormat[inproceedings]{doi}{}
%\DeclareFieldFormat[article]{doi}{}
\DeclareFieldFormat[inproceedings]{editor}{}
\DeclareFieldFormat[proceedings]{editor}{}
\DeclareFieldFormat[article]{editor}{}
\DeclareFieldFormat[inproceedings]{isbn}{}
\DeclareFieldFormat[proceedings]{isbn}{}
\DeclareFieldFormat[article]{isbn}{}

@@

%\vspace{-1em}
\section{Introduction}
\begin{figure}
	\centering
		\includegraphics[keepaspectratio, width=0.8\linewidth]{figures/arch-with-p4}
		\caption{\approachshort{} brings low-latency, online reinforcement learning to SoC- and NPU-based SmartNICs. Classical RL policy methods are the key to making this computationally feasible.\label{fig:netro-arch}}
\end{figure}
\begin{figure}
	\centering
%	\includegraphics[keepaspectratio, width=0.6875\linewidth]{figures/arch-with-p4}
%	\caption{\approachshort{} brings low-latency, online reinforcement learning directly to the dataplane. SoC- and NetFPGA-based SmartNIC devices expose spare compute---making in-situ, asynchronous processing and learning possible alongside P4 dataplanes. Classical RL policy methods are the key to making this computationally feasible.\label{fig:netro-arch}}
	\centering
	\includegraphics[keepaspectratio, width=0.85\linewidth]{figures/coop}
	\caption{Parallel policy execution. A single \emph{controller} delegates RL computation and updates to many \emph{minion} threads, using a shared atomic writeback.\label{fig:single-and-parallel:parallel}}
\end{figure}
Automatic network optimisation, control, and defence are becoming commonplace through data-driven techniques such as RL methods, where every change and its effects improve future decisions made by an agent.
In tandem, P4~\parencite{DBLP:journals/ccr/BosshartDGIMRSTVVW14} and \emph{programmable dataplane} hardware have inspired explosive growth and interest in the research community surrounding in-network computation and offloading.
This ecosystem exposes not only runtime reconfigurable packet processing, but per-packet and per-flow traffic measurement state that can greatly aid network operation~\parencite{DBLP:conf/sosr/GhasemiBR17}.

@@

%?? Do we want concrete examples of things `pushed down the stack'?
%
%?? Consider a DDoS prevention system...
%?? in the face of new data, Able to start learning again in microseconds
\fakepara{Machine learning in the dataplane}
To process this fine-grained state both at line rate \emph{and} at low latency, there has been keen interest in executing ML models in the dataplane~\parencite{DBLP:conf/hotnets/XiongZ19,DBLP:journals/corr/abs-2009-02353}.
Works in this category, while powerful, operate by converting a \emph{pre-trained} model into a suitable representation, such as a binary neural network or string of match-action tables.
%These works have shown the value of in-network ML: high-throughput, low latency response to network changes.
These can react to on-device state, yet the missing piece of the puzzle is learning and updating these ML analyses \emph{online}, without deferring to another machine in the network.
%Training these models online and in-network is an exciting (and challenging) lacuna in the field that \emph{has yet to be addressed by the community}.
This lacuna has yet to be addressed by the community.

\fakepara{Problems of online training}
Programmable dataplane hardware, being designed solely for efficient packet processing, lacks floating-point arithmetic support, even in the case of more general purpose NPU-type SmartNICs.
Additionally, deep neural network training relies on backpropagation, can require memorizing many a sizeable replay buffer for RL, and needs many mini-batches of data for stable training.
The natural solution is to pass training data to a host machine or dedicated accelerator, yet this adds delays in crossing the PCIe bus, kernel-user handover, and buffering---adding latency while learning.
The above (NIC-suitable) inference schemes still must be trained offline, even though their data formats solve the issue of floating point use.
Online learning thus requires careful choices in sample data formats and the learned function approximation scheme.

\section{\approachshort{}'s design}
\approachshort{} makes RL workloads feasible on SmartNICs by using quantised fixed-point representations of action values, tile-coded policies, and one-step temporal-difference RL algorithms such as Sarsa~\parencite{RL2E}.
These allow us to evaluate and update policies using \emph{only integer arithmetic}, and are computationally simple.
We observe that tile coded inference is a map-reduce problem (\cref{fig:par-tilecode}), and thus over integers admits a novel \emph{wait-free Sarsa} RL algorithm (exploiting the parallelism built in to SmartNICs).
While these functions have lower theoretical model capacity, they do not require batches of inputs to learn in a stable way, negating the memory needed to store experience replays or minibatches.

@@

\fakepara{Inference/Updates}
State is sent to a set of packet processing threads, which compute hit tiles from a locally held work set.
During inference, these threads atomically add values to a shared action preference list, while tile-coding guarantees that policy updates proceed without data races (\cref{fig:single-and-parallel:parallel}).
In the case of our Netronome implementation, these run on lightweight cores known as \emph{microengines} (MEs)---which could be substituted for suitable FUs on CGRA/FPGA hardware.
Additionally, we have a single-threaded implementation, \indfw, which forgoes this cooperative model to maximise offline throughput (by having many parallel single-threaded agents).

@@

\section{Evaluation}
We compare our implementation of \approachshort{} against a tile-coded RL agent written in numpy on a powerful host (i7-6700K@\qtyproduct[product-units=single]{4 x 4.2}{\giga\hertz}), using the policy dimensions of existing DDoS prevention work~\parencite{DBLP:journals/tnsm/SimpsonRP20}, across various quantisation bit depths.
\Cref{fig:lat-cumul} shows that our main design (\coopfw) achieves order-of-magnitude improvements over our single-threaded variant (\indfw) and a host implementation, as well as considerably tighter tail behaviour on the SmartNIC.
Note that lower bit depths have higher execution costs due to the native \qty{32}{\bit} width of Netronome registers.
On update times, we see \qty{63.12}{\micro\second} at 99.99\nthscript{th} percentile for a \qty{32}{\bit} policy of this size.
We relate a \qty{15}{$\times$} latency (\coopfw) and \qty{2.82}{$\times$} offline throughput-per-core (\indfw) in spite of the Netronome's considerably slower clock speed (\qty{0.29}{$\times$}), showing the value of exploiting SmartNIC parallelism.
